---
title: "Rising Sea level effects on Land Masses"
author: "Bipul"
---

```{r}
knitr::opts_chunk$set(warning = FALSE)
```

```{r}
library(ggplot2)
```

## **1. The data**

We check the present directory for the files that are present and then we will import the **datasets**.

```{r}
dir()
```

Now we import the **datasets** that will be required for this problem and store them in 2 variables, **sea_data** and **sea_level**

```{r}
sea_data <- read.csv("sea_data.csv")
sea_level <- read.csv("sea-level.csv")
```

We take a look at the summary of the 2 datasets:

```{r}
str(sea_data)
str(sea_level)
```

## **2.Analysis**

### 2.1 Sea Level dataset:

We will analyse this dataset to check the **general trends** of the **rising sea level**.  

First we change the names of the columns of the **sea_level** dataset.
```{r}
names(sea_level) <- c("Year", "Sea Level(inches)", "Lower Bound Level", "Upper Bound Level", "Sattelite")
```

Now we make the plots for the **Years** VS **Sea Level**.

```{r}
s <- ggplot(sea_level, aes(Year, `Sea Level(inches)`+ (`Lower Bound Level`+`Upper Bound Level`)/2))
s + geom_line(color = c("black")) + geom_smooth(linetype = 3, size = 2, color = "blue") + labs(y = "Adjusted Sea Level(incles)", title = "Sea Level Rise VS Year(1880 to 2014)") + theme_classic()
```

The graph shows us an overall increase in the sea level every year. The **blue regression line** shows a **general increasing trend** in the **sea level** over the **years**, the **grey** area around the line indicates the **95% Confidence Interval**.

### 2.2 Sea and Land dataset:

Now we work with the **sea_data** dataset.

#### **2.2.1 Countries VS Total Land Mass**

```{r}
sub_sa_af <- subset(sea_data, Region == "Sub Saharan Africa")
not_sub_saharan <- subset(sea_data, Region != "Sub Saharan Africa")
```

```{r}
dim(sub_sa_af)
dim(not_sub_saharan)
```

So we have separated all the data for each region in separate variables.
Now we will make plots for each region.

```{r}
qplot(Total.Area.sq.m., Country.Name, data = sub_sa_af, xlab = "Total Area in sq. meters", main = "Sub Saharan Africa")
qplot(Total.Area.sq.m., Country.Name, data = not_sub_saharan , xlab = "Total Area in sq. meters", main = "Other Countries as per their region", color = Region)
```

#### **2.2.2 Submerged area with rising sea level**

##### First we see the plots for the **Sub Saharan Africa** region.
**a.What effect does a rise in 1 meter sea level has on the land masses?**
```{r}
g <- ggplot(sub_sa_af, aes( X..affected.area.by.1m.rise, Country.Name))
g + geom_point() + labs(x = "% of land submerged", y = "Country Name", title = "Land Submerged when Sea level rises by 1m")
```

**b.What effect does a rise in 2 meter sea level has on the land masses?**
```{r}
g <- ggplot(sub_sa_af, aes( X..affected.area.by.2m.rise, Country.Name))
g + geom_point() + labs(x = "% of land submerged", y = "Country Name", title = "Land Submerged when Sea level rises by 2m")
```

**c.What effect does a rise in 3 meter sea level has on the land masses?**
```{r}
g <- ggplot(sub_sa_af, aes( X..affected.area.by.3m.rise, Country.Name))
g + geom_point() + labs(x = "% of land submerged", y = "Country Name", title = "Land Submerged when Sea level rises by 3m")
```

**d.What effect does a rise in 4 meter sea level has on the land masses?**
```{r}
g <- ggplot(sub_sa_af, aes( X..affected.area.by.4m.rise, Country.Name))
g + geom_point() + labs(x = "% of land submerged", y = "Country Name", title = "Land Submerged when Sea level rises by 4m")
```

**e.What effect does a rise in 5 meter sea level has on the land masses?**
```{r}
g <- ggplot(sub_sa_af, aes( X..affected.area.by.5m.rise, Country.Name))
g + geom_point() + labs(x = "% of land submerged", y = "Country Name", title = "Land Submerged when Sea level rises by 5m")
```

##### Now we see the plots for the **countries other than the Sub Saharan Africa** region.
**a.What effect does a rise in 1 meter sea level has on the land masses?**
```{r}
g <- ggplot(not_sub_saharan, aes( X..affected.area.by.1m.rise, Country.Name, color = Region))
g + geom_point() + labs(x = "% of land submerged", y = "Country Name", title = "Land Submerged when Sea level rises by 1m")
```

**b.What effect does a rise in 2 meter sea level has on the land masses?**
```{r}
g <- ggplot(not_sub_saharan, aes( X..affected.area.by.2m.rise, Country.Name, color = Region))
g + geom_point() + labs(x = "% of land submerged",y = "Country Name", title = "Land Submerged when Sea level rises by 2m")
```

**c.What effect does a rise in 3 meter sea level has on the land masses?**
```{r}
g <- ggplot(not_sub_saharan, aes( X..affected.area.by.3m.rise, Country.Name, color = Region))
g + geom_point() + labs(x = "% of land submerged", y = "Country Name", title = "Land Submerged when Sea level rises by 3m")
```

**d.What effect does a rise in 4 meter sea level has on the land masses?**
```{r}
g <- ggplot(not_sub_saharan, aes( X..affected.area.by.4m.rise, Country.Name, color = Region))
g + geom_point() + labs(x = "% of land submerged", y = "Country Name", title = "Land Submerged when Sea level rises by 4m")
```

**e.What effect does a rise in 5 meter sea level has on the land masses?**
```{r}
g <- ggplot(not_sub_saharan, aes( X..affected.area.by.5m.rise, Country.Name, color = Region))
g + geom_point() + labs(x = "% of land submerged", y = "Country Name", title = "Land Submerged when Sea level rises by 5m")
```


## **3. Work Flow**

a. An **Exploratory Analysis** on the provided dataset was done.
b. The main aim of the Analysis was to **understand the effects of rising sea level on the land masses**.
c. We import the 2 datasets and do a bit of **pre processing**
d. The first data set shows the **globally rising sea level**.
e. We then look at the **total area of land masses in each country**. The countries were splitted into 2 sections **because plotting all the countries in one data set would have made the plot very clumsy!!**
f. Now, we make **5 different plots** for **Sub Saharan Africa** countries and the **area that will be submerged** if the **sea level rises by 1, 2, 3, 4, 5 meters**
g. Now, we make **5 different plots** for **non - Sub Saharan Africa** countries and the **area that will be submerged** if the **sea level rises by 1, 2, 3, 4, 5 meters**

## **4. Conclusion**

a. In the first plot we obsereve that the **global sea level is rising**. This is shown by the **blue regression line**.
b. The same plot also demonstrates the **nature of increasing sea level is exponential**.
c. The plots related to the **Land % submerged VS sea level** depicts a picture in front of us about the effect a rising sea level is going to have on the land masses.
d. The plots for **other countries are splitted as per the Regions**, for an overview of the **% of land that will be under water if the sea level rises by X%**
e. The Analysis was effected because, even after a lot of internet browsing, **an appropriate all in one data set was not found!!**
f. Sites like **GitHub** and **Kaggle** were browsed to find a suitable dataset!!